---
title: "Hands on"
author: "Che Hidayat Che Nawi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(haven)
stroke <- read.csv("rconf.csv", stringsAsFactors = TRUE)

stroke$age <- as.numeric(stroke$age)

summary(stroke)

library(lubridate)
library(gtsummary)

stroke$dodiag <- as.Date(stroke$dodiag)
stroke$dodeath <- as.Date(stroke$dodeath)
stroke$dodis <- as.Date(stroke$dodis)
stroke <- stroke %>% mutate(dur = stroke$dodiag %--% stroke$dodeath) %>% 
  mutate(dur = as.duration(dur))

stroke <- stroke %>% mutate(dur_days = dur/ddays(1))

stroke <- stroke %>% mutate(dur_month = dur/ddays(1)/30.417)

str(stroke)

stroke2 <- stroke %>% select(age, sex, dm, hpt, nihss, status_f.u, dur_days,dur_month)
stroke2_complete <- na.omit(stroke2)
stroke2_complete$nihss <- relevel(stroke2_complete$nihss, ref = "No stroke symptoms (0)")
```

# Application and examples

## Descriptive analysis

1. tbl_summary()

```{r}
library(gtsummary)
tbl_summary(stroke2_complete)
```

2. tbl_cross()

```{r}
tbl_cross(stroke2_complete,
          row = sex,
          col = status_f.u,
          percent = "row",
          margin = "row") %>% 
  add_p(source_note = TRUE)
```

## Binary logistic regression

### Default output for regression analysis in R

```{r}
default_stroke <- glm(status_f.u~age + sex, data = stroke2_complete, family = binomial(link = logit))

summary(default_stroke)
```

### Logistic regression using gtsummary-univariate analysis

```{r}
uvlog <- stroke2_complete %>% select(age, sex, dm, hpt, nihss, status_f.u) %>% 
  tbl_uvregression(method = glm,
                   y=status_f.u,
                   method.args = list(family = binomial),
                   exponentiate = TRUE)

uvlog
```

### Logistic regression using gtsummary-multivariate analysis

```{r}
mvlog <- glm(status_f.u~age+sex+dm+hpt+nihss,
             stroke2_complete, 
             family = binomial)
mvlog %>% tbl_regression(
  exponentiate=TRUE
)
```

## Survival analysis

### Survival rate using tbl_survfit

```{r}
library(survival)
library(gtsummary)
fit1 <- survfit(Surv(dur_month, status_f.u)~ 1, stroke2_complete)
fit2 <- survfit(Surv(dur_month, status_f.u)~ sex, stroke2_complete)

life_table <- list(fit1, fit2) %>% 
  tbl_survfit(times= c(1, 12, 36)) %>% 
  modify_header(update = list(
    stat_1 ~ "**1-month**",
    stat_2 ~ "**1-year**",
    stat_3 ~ "**3-years**"
  )) %>% 
  add_n() 

life_table
```


### Semi-parametric survival using gtsummary-univariate analysis

```{r}
library(survival)
cox_uv <- tbl_uvregression(
  stroke2_complete[c("dur_month", "status_f.u", "age", "sex", "dm",
             "hpt", "nihss")],
  method = coxph,
  y = Surv(time = dur_month, event = status_f.u=='Died'),
  exponentiate = TRUE
)

cox_uv
```

### Semi-parametric survival using gtsummary-multivariate analysis

```{r}
cox_mv <- coxph(Surv(time = dur_month, event = status_f.u=='Died')~
                  age+sex+dm+hpt+nihss,
             stroke2_complete) %>% 
  tbl_regression(exponentiate=TRUE)
  
cox_mv
```

# Customization

## Data arrangement

```{r}
stroke3 <-  stroke2_complete %>% select(age, sex, dm, hpt, nihss)
desc <- tbl_summary(stroke3,
            by    = sex,
            label = list( age                 ~    "Age",
                          sex                 ~    "Gender",
                          dm                  ~    "Diabetes Mellitus",
                          hpt                 ~    "Hypertention",
                          nihss               ~    "NIHSS Score"),
            digits =      c(all_continuous()  ~    1,
                            all_categorical() ~    0),
            statistic =   c(all_categorical() ~    "{n} ({p}%)",
                            all_continuous()  ~    "{mean} ({sd})"))

desc 
```

## Add extra information

```{r}
#for descriptive table
desc %>% 
  add_n %>% 
  add_p %>% 
  add_q 
```

## Aesthethic

```{r}
desc %>% 
  add_n %>% 
  add_p %>% 
  add_q %>% 
  bold_labels() %>% 
  italicize_levels()

#may add bold_p if there is significant differenc with p value of <0.05
```

# Merging and stacking


## Merging

```{r}
cox_uv <- tbl_uvregression(
  stroke2_complete[c("dur_month", "status_f.u", "age", "sex", "dm",
             "hpt", "nihss")],
  method = coxph,
  y = Surv(time = dur_month, event = status_f.u=='Died'),
  exponentiate = TRUE,
  label = list( age                 ~    "Age",
                          sex                 ~    "Gender",
                          dm                  ~    "Diabetes Mellitus",
                          hpt                 ~    "Hypertention",
                          nihss               ~    "NIHSS Score"),
)

```

```{r}
cox_mv <- coxph(Surv(time = dur_month, event = status_f.u=='Died')~
                  age+sex+dm+hpt+nihss,
             stroke2_complete) %>% 
  tbl_regression(exponentiate=TRUE,
                 label = list( age                 ~    "Age",
                          sex                 ~    "Gender",
                          dm                  ~    "Diabetes Mellitus",
                          hpt                 ~    "Hypertention",
                          nihss               ~    "NIHSS Score"),)
  
```

```{r}
tbl_surv_merge <- tbl_merge(
  list(cox_uv, cox_mv),
  tab_spanner = c("**Univariable**","**Multivariable**")
)
tbl_surv_merge
```

## Stacking

```{r}
t1 <- glm(status_f.u~sex,
          data = stroke2_complete,
          family = binomial) %>% 
  tbl_regression(
    exponentiate=TRUE,
    label=list(sex ~"Gender (unadjusted)")
  )

t2 <- glm(status_f.u~sex+age+dm+hpt,
          data = stroke2_complete,
          family = binomial) %>% 
  tbl_regression(
    include="sex",
    exponentiate=TRUE,
    label=list(sex ~"Gender (adjusted)")
  )

table_stack_ex1 <- tbl_stack(list(t1, t2))
table_stack_ex1
```

# Print engine (gtsummary + R Markdown)


```{r}
library(flextable)
tbl_surv_merge %>% 
  bold_labels() %>% 
  italicize_levels() %>%
  as_flex_table() # if knit to microsoft word
```

# Journal table format

```{r}
reset_gtsummary_theme()
theme_gtsummary_journal(journal = "lancet")

lancet_theme <- tbl_surv_merge %>% 
  bold_labels() %>% 
  italicize_levels() %>%
  as_gt() %>% 
  gt::tab_header("Journal Theme (Lancet)")

lancet_theme
```

# Report statistics in line

Tables are important but sometimes we still need to report a result in-line in a report. This is especially true when explaining the results of regression models for reader’s understanding.

With the inline_text() function, any data reported in the tables using {gtsummary} can be extracted and reported in-line in R markdown

For example:

We want to report hazard ratio for age from table cox_mv

```{r}
cox_mv
```

For every 1 year increment in age, the hazard of dying from stroke increase by 4%, `r inline_text(cox_mv, variable = age)`.
